---
title: "Network-based for linear models in R"
#author: "Zeus Gracia-Tabuenca"
#date: "March 2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NBR-LM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette documents the implementation of NBR 0.1.2 for linear models applied to a sample level of matrices edges.

Here it is an example applied into a 3D volume. Where, its first two dimension must be equal, i.e., the number of nodes of every individual network matrix (`nnodes`). And its third dimension equals the sample number.

```{r setup}
library(NBR)
cmx <- NBR:::frontal3D          # Load 3D array
brain_labs <- NBR:::frontal_roi # Load network nodes labels
phen <- NBR:::frontal_phen      # Load sample phenotypic info
dim(cmx)                        # Show 3D array dimensions
```

We can plot the sample average network matrix, with `lattice::levelplot`.

```{r input networks, fig.align = "center"}
library(lattice)
avg_mx <- apply(cmx, 1:2, mean)
# Set max-absolute value in order to set a color range centered in zero.
flim <- max(abs(avg_mx)[is.finite(avg_mx)])
levelplot(avg_mx, main = "Average", ylab = "ROI", xlab = "ROI",
          at = seq(-flim, flim, length.out = 100))
```

As we can observe, this is a symmetric matrix with the pairwise connections of the 28 regions of interest (ROI) `brain_labs`. The next step is to check the sample inference data to be tested edgewise, in this case `phen`. Before applied the NBR-LM it is necessary to be sure that the number of networks (i.e., the third dimension of the 3D array) matches the number of observations of the data.frame one wants to test.

```{r input phenotypic info}
head(phen)
nrow(phen)
identical(nrow(phen), dim(cmx)[3])
```

The data.frame contains the individual information for diagnostic group, sex, and chronological age. So, there are all the ingredients to perform a NBR-LM. We are going to test first a group inference (with just ten permutations in one core).

```{r group-based NBR}
set.seed(18900217) # Because R. Fisher is my hero
before <- Sys.time()
nbr_group <- nbr_lm_aov(net = cmx, nnodes = 28, idata = phen,
   mod = "~ Group", thrP = 0.01, nperm = 10)
after <- Sys.time()
show(after-before)
```

Although ten permutations is faraway for a proper null distribution, we can see that it takes several seconds to performed it. So we suggest to parallel to multiple CPU cores with `cores` argument.

```{r multicore group-based NBR, eval = FALSE}
set.seed(18900217)
library(parallel)
before <- Sys.time()
nbr_group <- nbr_lm_aov(net = cmx, nnodes = 28, idata = phen,
   mod = "~ Group", thrP = 0.01, nperm = 100, cores = detectCores())
after <- Sys.time()
length(nbr_group)
```

NBR functions return a nested list of at least two list. The first list encompasses all the individual significant edges and its corresponding component and statistical strength above threshold (p < 0.01, in this example). In this case all the significant edges belong to a single component.

```{r component display, fig.align = "center"}
# Plot significant edges
edge_mat <- array(0, dim(avg_mx))
edge_mat[nbr_group$components$Group[,2:3]] <- 1
levelplot(edge_mat, col.regions = rev(heat.colors(100)),
          main = "Component", ylab = "ROI", xlab = "ROI")
show(nbr_group$fwe$Group)
```

As we can observe, significant edges are display in the upper triangle of the matrix, and the second list `fwe` contains, for each term of the equation, the probability of the observed values within the generated null distribution.
